---
title: "Using `smartadata`"
author: "Ignacio Cordón"
date: "`r Sys.Date()`"
output: rmarkdown::pdf_document
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE, 
  message = FALSE
)
```


## Purpose of `smartdata`
R programming language has a wide variety of packages that target machine learning topiscs. However, each package has its own interface, which makes the task of using several of them quite time-consuming. The main purpose of `smartdata` is to provide a common interface for a collection of well-used machine learning packages, so that using methods from different libraries gets easier. Also, it standardizes names of arguments, so that if a method had a parameter `num_iterations` and other had a parameter `iterations` which had a similar meaning, both methods have the same name for the parameter in `smartdata`.

`smartdata` includes preprocessing algorithms for oversampling, instance selection, feature selection, normalization, discretization, space transformation, outliers treatment, missing values treatment and noise cleaning.

Each of the aforementioned topics has its corresponding wrapper: `instance_selection`, `feature_selection`, `normalize`, `discretize`, `space_transformation`, `clean_outliers`, `impute_missing` and `clean_noise`.

In addition to that, `magrittr` has been used to provide a pipeline, so that a workflow can be expressed as:

```{r, eval = FALSE}
result <- dataset %>% impute_missing %>% clean_noise %>% oversample %>% feature_selection
```

## Basic help
To check methods a certain wrapper can be called with, we can simply call the `help` function included in `smartdata`: `which_options`.

```{r, warning = FALSE, message = FALSE}
library("smartdata")
which_options("instance_selection")
```

To check parameters for a wrapper, it suffices to do `?{name of the wrapper}`. For example: `?instance_selection`, which would output:

```
Usage

instance_selection(dataset, method, class_attr = "Class", ...)

Arguments

dataset	    we want to perform an instance selection on
method	    selected method of instance selection
class_attr	character. Indicates the class attribute from dataset. Must exist in it
...         Further arguments for method
```

Moreover, if we already know which method we want to use but we do not recall its arguments, we can check its list of parameters with:

```{r}
which_options("instance_selection", "multiedit")
```

That option provides a brief description for each possible parameter, as well as a reference to the original function, in case the information for some parameter might not seem clear enough (although mapping between original function's arguments and `smartdata` wrapper is not exact).

In summary, a valid call for `multiedit` would be:
```{r}
super_iris <- iris %>% instance_selection("multiedit", k = 3, num_folds = 2, 
                                          null_passes = 10, class_attr = "Species")
```

Or:
```{r}
super_iris <- iris %>% instance_selection("multiedit", k = 3, null_passes = 10,                                           
                                          class_attr = "Species")
```

Or:
```{r}
super_iris <- iris %>% instance_selection("multiedit", k = 3, 
                                          class_attr = "Species")
```

Or even:
```{r}
super_iris <- iris %>% instance_selection("multiedit", class_attr = "Species")
```

# Techniques included in `smartdata`
Let $S = \{(x_i, y_i)\}_{i=1^m}$ be our set of data from this moment on. It holds $x_i \in \mathcal{X} \subset \mathbb{R}^n$ and $y_i\in \mathcal{Y}$ where $\mathcal{Y}$ is a finite set of labels. $(x_i, y_i)$ is called instance.

## Oversampling
Oversampling consists in given $\mathcal{Y}$ with $|\mathcal{Y}| = 2$, and more instances labeled with one class than with the other, generating synthetic instances labeled with the minority class, namely $E$ so that the resulting dataset $S \cup E$ has better characteristics for classification. 

Since oversampling is going to replicate or generate artificial instances belonging to the minority class, a class attribute must be indicated with `class_attr` argument, as well as a `ratio` of a desired ratio of imbalancement (when applicable), where this ratio is computed as:
\[
\frac{\textrm{number of minority instances}}{\textrm{number of majority instances}}
\]

Also, a `filtering` argument can be provided, indicating wHether to to perform a filtering of the generated instances, using NEATER.

Possible methods are:

* PDFOS
* RWO
* ADASYN
* ANSMOTE
* SMOTE
* MWMOTE
* BLSMOTE
* DBSMOTE
* SLMOTE
* RSLSMOTE
* RACOG
* wRACOG 

Only if picked method is wRACOG, an additional parameter `wrapper` with possible values 'KNN' and 'C5.0' can be provided. That parameter indicates the desired classificator to select instances.

As example:
```{r, message = FALSE, warning = FALSE, results = "hide"}
data(iris0, package = "imbalance")
super_iris <- oversample(iris0, method = "MWMOTE", class_attr = "Class",
                         ratio = 0.8, filtering = TRUE)
```

## Instance selection
Consists in picking a subset $S'\subseteq S$, so that certain characteristics are preserved with respect to the original sample (such as distribution of classes) or at least we keep most of representative instances.

As said, methods try to pick instances preserving original classes distribution, so `class_attr` must be supplied.

Possible methods are:

* CNN
* ENN
* multiedit
* FRIS

As example:
```{r, message = FALSE, warning = FALSE, results = "hide"}
super_iris <- instance_selection(iris, method = "CNN", class_attr = "Species")
```

## Feature selection
Consists in given a subset of features $T = \{i: i \in \{1,\ldots n\}$, projecting the tuples of $S = \{(x_i, y_i)\}_{i=1}^m$ to the set of features given by $T$. That is, if we defined 
\[
p_{T}((x_{i1}, \ldots, x_{in})) = (x_{ij})_{j \in T}
\]
as the projection to the features in $T$, doing a feature selection would result in a set $S' = \{(p_T(x_i), y_i)\}_{i=1}^m$, if the picked features were $T$.

Appart from the parameters for each specific method, `class_attr` must be supplied for each of the preprocessings, and an additional parameter `exclude` can be supplied with a vector of features names, so that those features are striped before the feature selection and joined after the procedure.

Possible methods are:

* Boruta
* chi_squared
* information_gain
* gain_ratio
* sym_uncertainty
* oneR
* RF_importance
* best_first_search
* forward_search
* backward_search
* hill_climbing
* cfs
* consistency 

## Normalization
Implies converting the sample data $S$ into another dataset $S'$ where each tuple is treated so that standard deviation of the data ends up being zero (feature-wise or considered as elements of $\mathbb{R}^n$), or all the data is mapped to $[0,1]$ interval (e.g. dividing by the maximum feature-wise), etc

Possible methods are:

* z_score
* pos_standardization
* unitization
* pos_unitization
* min_max
* rnorm
* rpnorm
* sd_quotient
* mad_quotient
* range_quotient
* max_quotient
* mean_quotient
* median_quotient
* sum_quotient
* ssq_quotient
* norm
* pnorm
* znorm
* decimal_scaling
* sigmoidal
* softmax

Appart from the parameters for each specific method, an additional parameter `exclude` can be supplied with a vector of features names, so that those features are striped before the normalization and joined after the procedure, with the unchanged original data.

As an example:

```{r, results = "hide", warning = FALSE, message = FALSE}
super_iris <- normalize(iris, method = "min_max", exclude = "Species", by = "column")
```
